\documentclass[aspectratio=169]{beamer}
\usetheme{default}
\usepackage{booktabs}
\definecolor{editblue}{RGB}{38,117,168}
\setbeamercolor{structure}{fg=editblue}
\title{How much edit experience should the model receive?}
\subtitle{Sequence-edit JEPA validity audit and representation decision}
\date{17 July 2026}
\begin{document}
\frame{\titlepage}

\begin{frame}{The task is synthetic typo repair}
  \begin{itemize}
    \item Official iGSM supplies a problem and its clean, multi-step worked solution.
    \item We damage that solution with token insertions, deletions, and replacements.
    \item The expert path is the exact reverse of those corruptions, so it is oracle denoising rather than naturally observed editing.
    \item More independent damaged solutions add breadth; more alternative edits per state add local action coverage.
  \end{itemize}
  \vfill
  \textbf{Why it matters:} we must establish healthy primitive dynamics before asking whether hierarchy helps.
\end{frame}

\begin{frame}{The old hierarchy round is invalid}
  \begin{itemize}
    \item All five jobs failed before their first optimizer step, so they provide no architectural evidence.
    \item The immediate symptom was a 276-token ``sentence'' exceeding a 128-position table.
    \item The cause was deeper: official steps often end in fused tokens such as \texttt{6.}, while the adapter split only on standalone \texttt{.}.
    \item The corrected adapter preserves official nested steps and exactly reverses every audited corruption.
  \end{itemize}
  \vfill
  \textbf{What to notice:} this is an invalid result, not evidence against hierarchy.
\end{frame}

\begin{frame}{The corrected data contract passes its first gate}
  \centering
  \begin{tabular}{lr}
    \toprule
    Audit check & Observed result \\
    \midrule
    Exact terminal recovery & 128/128 \\
    Multi-step solution collapse & 0/128 \\
    Official solution length & median 166.5 tokens \\
    Official steps & median 9 \\
    Maximum step length & p95 43 tokens \\
    Repairs per trajectory & mean 11.14 \\
    Repair/minimum-distance ratio & mean 1.029 \\
    \bottomrule
  \end{tabular}
  \vfill
  \textbf{Why it matters:} the model now sees the intended step sequence, but these are data-validity observations, not trained-model results.
\end{frame}

\begin{frame}{All six process-valid cells ignore the edit action}
  \centering
  \begin{tabular}{lrrr}
    \toprule
    Cell & Prediction & No change & Shuffled/matched \\
    \midrule
    512 trajectories & .397 & .575 & 1.00014 \\
    2,000 trajectories & .394 & .302 & 1.00000 \\
    6,000 trajectories & .224 & .113 & 1.00005 \\
    K=1 at 2,000 & .287 & .268 & 1.00003 \\
    K=4 at 2,000 & .175 & .123 & 1.00003 \\
    \bottomrule
  \end{tabular}
  \vfill
  \textbf{What to notice:} lower raw error does not establish dynamics; most cells lose to copying the current state, and shuffling the action changes nothing.
\end{frame}

\begin{frame}{One-token effects disappear in the global target}
  \begin{itemize}
    \item Four exact global next-buffer targets from one state are separated by only .000228 normalized L1.
    \item The exact changed-step targets for those edits are separated by .634, about 2,775 times more.
    \item The old K=4 model assigns changed-step outcomes at 24.7\%, chance for four alternatives.
    \item The repair predicts exact changed-step anchors for expert and mechanical alternative edits, with no quality or preference labels.
    \item Coefficients .25, 1, and 4 plus a lower-learning-rate check all completed cleanly.
  \end{itemize}
  \vfill
  \textbf{Question:} does stronger local supervision make the pooled predictor use the action?
\end{frame}

\begin{frame}{Local supervision still leaves the model action-blind}
  \centering
  \begin{tabular}{lrrrr}
    \toprule
    Weight & Pred. & Persist. & Shuf./match & Assignment \\
    \midrule
    .25 & .173 & .130 & 1.00004 & 24.92\% \\
    1 & .173 & .146 & 1.00003 & 24.85\% \\
    4 & .186 & .150 & 1.00151 & 25.04\% \\
    4, low LR & .354 & .177 & 1.00041 & 25.00\% \\
    \bottomrule
  \end{tabular}
  \vfill
  \begin{itemize}
    \item Assignment stays at 25\% chance; every cell loses to persistence.
    \item Rank falls 11.6--26.4\%; no coefficient deserves confirmation.
    \item The ``local'' head still projects one pooled global-state prediction.
  \end{itemize}
  \textbf{Next decision:} one direct observed-step attention test under the unchanged gates; otherwise retire this state family.
\end{frame}
\end{document}
