\documentclass[aspectratio=169]{beamer}
\usetheme{Madrid}
\usepackage{booktabs}
\usepackage{graphicx}
\title{Making a learned intent catalogue executable}
\subtitle{Token-level prerequisite support, 21 July 2026}
\author{TextJEPA research cycle}
\date{}

\begin{document}
\maketitle

\begin{frame}{The everyday problem: do I have the prerequisite?}
  \begin{itemize}
    \item The controller sees a shuffled box of instruction cards called intent phrases.
    \item An instruction is executable only after its prerequisite computations have run.
    \item Compressing each earlier instruction into one vector can blur the exact entity names.
    \item We therefore let candidate words attend directly to words in the executed intent history.
    \item Success here means proposing executable actions without consulting a hidden feasible-action menu.
  \end{itemize}
\end{frame}

\begin{frame}{Token history repairs the proposal interface}
  \centering
  \begin{tabular}{lrrr}
    \toprule
    Controller & Strict & Plus two & Invalid\\
    \midrule
    Token head, history masked & .025 & .025 & .950\\
    Phrase-pooled history & .083 & .358 & .483\\
    Token-level history & \textbf{.175} & \textbf{.717} & \textbf{.158}\\
    \bottomrule
  \end{tabular}
  \vspace{1em}

  \begin{minipage}{0.9\linewidth}
  \small What to notice: token history roughly doubles the best observed
  phrase-pooled length-nine success and cuts invalid attempts by about two
  thirds. Why it matters: JEPA simulation can now be tested behind a viable,
  non-oracle proposal interface.
  \end{minipage}
\end{frame}

\begin{frame}{The scientific boundary is essential}
  \begin{itemize}
    \item These are 120-episode, one-training-seed validation results.
    \item Every controller sees the same complete, shuffled, outcome-free intent catalogue.
    \item The negative control has equal token-head capacity but cannot see executed intent tokens.
    \item Selection was prior-only: JEPA consequence reranking had weight zero.
    \item The result supports lexical prerequisite tracking, not a claim that latent simulation improved planning.
  \end{itemize}
\end{frame}

\begin{frame}{Next decision: does simulation add value after support is fixed?}
  \begin{itemize}
    \item Freeze the learning-rate-$10^{-3}$ token-support checkpoint and the action catalogue.
    \item Compare prior-only selection with one-step and deeper causal JEPA reranking.
    \item Hold top-M, beam width, support weight, episodes, and problem order fixed.
    \item Scale only if reranking raises success without restoring invalid actions.
    \item If reranking hurts, separate value error from recursive world-model drift before retraining.
  \end{itemize}
\end{frame}

\end{document}
