\documentclass[aspectratio=169]{beamer}
\usetheme{default}
\usepackage{booktabs}
\definecolor{editblue}{RGB}{38,117,168}
\setbeamercolor{structure}{fg=editblue}
\title{Can representation spreading make edits causally visible?}
\subtitle{Token-aligned sequence-edit JEPA: curriculum VICReg screen}
\date{17 July 2026}
\begin{document}
\frame{\titlepage}

\begin{frame}{The editor must care which edit we request}
  \begin{itemize}
    \item The model receives a damaged worked solution and a literal insert, delete, or replace action.
    \item It predicts the next internal representation without reconstructing every output token.
    \item A collapsed model can look accurate while treating many actions as equivalent.
    \item Variance-Invariance-Covariance Regularization (VICReg) spreads internal states and reduces redundant features.
  \end{itemize}
  \vfill
  \textbf{Question:} does stronger VICReg make the transition predictor depend on the requested edit?
\end{frame}

\begin{frame}{Five cells differ only in VICReg strength}
  \begin{itemize}
    \item All cells use 2,000 curriculum-corrupted oracle trajectories, seed 0, three epochs, and zero dropout.
    \item All cells share the token-pointer action interface, recursive depth-four predictor, and exponential-moving-average target.
    \item Tested whole-loss weights are 0, .02, .1, .5, and 1.0.
    \item The causal test swaps actions across valid states; error must rise by at least 5\%.
    \item Effective-rank loss must remain within 10\% of the healthiest control.
  \end{itemize}
  \vfill
  \textbf{Why it matters:} hierarchy remains blocked until primitive edit dynamics pass both gates.
\end{frame}

\begin{frame}{Every cell predicts, but every cell underuses the action}
  \centering
  \begin{tabular}{lrrrr}
    \toprule
    VICReg & One-step & No change & Shuf./match & Rank \\
    \midrule
    0 & .149 & .256 & 1.013 & 113.6 \\
    .02 & .153 & .263 & 1.013 & 107.4 \\
    .1 & .151 & .320 & 1.014 & 84.4 \\
    .5 & .149 & .346 & 1.010 & 70.2 \\
    1.0 & .160 & .272 & 1.009 & 69.5 \\
    \bottomrule
  \end{tabular}
  \vfill
  \begin{itemize}
    \item Every cell beats copying the current state, so optimization is functioning.
    \item No shuffled-action ratio approaches the 1.05 causal threshold.
    \item Weight .02 is the only nonzero setting within the rank-loss gate.
  \end{itemize}
\end{frame}

\begin{frame}{The next test directly supervises action geometry}
  \begin{itemize}
    \item Strong VICReg spreads values but removes 26--39\% of effective rank; it is not selected.
    \item Weight .02 remains only as a low-cost combination anchor, not as a successful result.
    \item Latent Difference Action Decoding (LDAD) reconstructs the complete observed edit phrase only from the online state displacement.
    \item The next screen tests LDAD weights 1, 10, and 20, plus no-VICReg and moderate-VICReg interaction controls at weight 10.
    \item LDAD continues only if it improves causal action use and recursive prediction, not merely action-token decoding.
  \end{itemize}
  \vfill
  \textbf{Limitation:} this remains one-seed, candidate-privileged oracle denoising, not autonomous reasoning.
\end{frame}
\end{document}
